### 2023-02-07 article.tex
Improve

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Making code useful paradata}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Code as paradata has multiple uses:
(1) to reproduce an experiment, by running the unchanged code on the same data as
used in a published research,
(2) to replicate an experiment, by running the unchanged code on different data as
used in a published research,
(3) to extend an experiment, by changing the code.

@@

For code to be ideal paradata, it must be documented.
However, code does explain itself to some extent,
for example, the function name \verb|t.test| as used above.

\begin{itemize}
  \item Code must be documented to the right amount
  \item Code must be reviewed
\end{itemize}

\paragraph{Code must be documented to the right amount}

Code should be documented to the right amount:
too little documentation is unhelpful, where too much is a needless burden
on the programmer.
The first recommended practice is to write the code in such a way 
that it becomes self-explanatory \cite{wilson2014best}.
What is left to document are the reasons behind the code, 
its design and its purpose \cite{wilson2014best}.

@@

There exist website that are dedicated to host code.
Such so-called code hosting websites have many advantages,
both for the authors and users of the code,
over copy-pasting code in the supplementary materials in a paper.

The practice of using a code hosting website has increased,
with the abstracts from articles in the academic journal
'Bioinformatics' referring to such a hosted website
has increased tenfold, going from below 5 percent in 2009 
to around 50 percent in 2017 \cite{russell2018large}.

\paragraph{Example of code hosting website}

There are multiple websites for hosting code, with BitBucket, GitHub,
GitLab and SourceForge being the most well-known.
GitHub is currenly the most popular website to host code in general,
with millions of users (a simple user search can confirm this).
GitHub users can create dedicated websites (called 'repositories')
to upload code. Code hosts, such as GitHub, 
store that uploaded code and provide ways 
for visiters to download the code and/or interact with it.
The use of such code hosting websites
accommodates collaboration \cite{perez2016ten}.
and improves transparency \cite{gorgolewski2016practical}.
An example from the author is the website \cite{bbbqarticleissue157},
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}
\iffalse
, as shown in Figure \ref{fig:bbbqarticleissue157}
\fi
.

@@

For a computational experiment to be reproducible,
an independent person should be able to
reproduce the same results.
The reproduce the same results, one needs the same
input data and the same code (and the same 
dependencies and environment).
When the input data are sensitive, however, and thus cannot be shared with
an independent researcher, one thus cannot reproduce the
experiment.

The solution for this problem is simple:
supply a simulated dataset with the code.
In the case of genetic epidemiology this would mean
simulated genotypes and associated phenotypes,
as can be done with the \verb|plinkr| R package \cite{plinkr}.
This solution, however, is known to all developer that follow
good practices when writing code, as public (i.e. simulated) 
datasets are needed for automatic testing of code.
One extra benefit of simulated data is that these can be used
as a benchmark, as slightly different analyses should give 
similar conclusions.

@@

Code needs maintanance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. 
The degradation of software is a known feature for nearly 
four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
here a more modern (and in this context superior)
definition is used: software collapse \cite{hinsen2019dealing},
in which software fails due to dependencies on other 
software.
Where that paper recommends using older software foundations 
with multiple implementations \cite{hinsen2019dealing},
this paper recommends a modern solution that potentially has more longevity:
to publish the code installed in a virtual environment
(including the dependencies used at that time), as described below.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Making code useful paradata}\label{sec:making-code-useful-paradata}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Code as paradata has multiple uses: to reproduce, to replicate and to 
extend \cite{benureau2018re}.

To reproduce an experiment is to obtain the same publish result,
by using the same code and data as used for that publication 
\cite{peng2006reproducible}.

(1) to reproduce an experiment, by running the unchanged code on the same data as
used in a published research,
(2) to replicate an experiment, by running the unchanged code on different data as
used in a published research,
(3) to extend an experiment, by changing the code.

@@

For code to be ideal paradata, it must be extensively documented.
This extend is determined by its intended use: 
to reproduce, replicate or extend \ref{sec:making-code-useful-paradata}.
The amount of documentation is, however, also determined
by the recommended practices for writing code in general.

\subsubsection{... for reproducing}

For reproducing purposes, the minimal amount of documentation
is determined by software development best practices.
When an independent researcher reproduces an experiment, 
it is possible that no code is read at all.
Due to this, one could argue that no documentation is needed.
However, all code should be documented 'adequately' \cite{peng2006reproducible},
ideally writing code in such a way that it becomes self-explanatory \cite{wilson2014best}
and for the remaining code to document the reasons behind it, 
its design and its purpose \cite{wilson2014best}.

\subsubsection{... for replication}

\subsubsection{... for extending}

@@

Code should be distributed in standard ways \cite{peng2006reproducible}.
The standard way to publish code is to use a code hosting website.
There are multiple code hosting websites, 
with GitHub being the most popular one \cite{cosentino2017systematic}.
GitHub is a website where users can create dedicated pages (called 'repositories')
to upload code and interact with it.
The use of such code hosting websites
has increased strongly \cite{russell2018large},
accommodates collaboration \cite{perez2016ten}
and improves transparency \cite{gorgolewski2016practical}.

\iffalse
An example from the author is the website \cite{bbbqarticleissue157},
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
as shown in Figure \ref{fig:bbbqarticleissue157}.
\fi

@@

Next to the code, it is the data used in an experiment 
that must be made available for an experiment to be called 'reproducible' \cite{peng2006reproducible}.
When the input data are sensitive, however, and thus cannot be shared with
an independent researcher, one thus cannot reproduce the
experiment.

The solution for this problem is simple:
supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
together with the code.
In the case of genetic epidemiology this would mean
simulated genotypes and associated phenotypes,
as can be done with the \verb|plinkr| R package \cite{plinkr}.
This solution, however, is known to all developer that follow
good practices when writing code, as public (i.e. simulated) 
datasets are needed for automatic testing of code.
One extra benefit of simulated data is that these can be used
as a benchmark, as slightly different analyses should give 
similar conclusions.

@@

Code needs maintanance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. 
The degradation of software is a known feature for nearly 
four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
here a more modern (and in this context superior)
definition is used: software collapse \cite{hinsen2019dealing},
in which software fails due to dependencies on other 
software.
Where that paper recommends using older software foundations 
with multiple implementations \cite{hinsen2019dealing},
this paper recommends a modern solution that potentially has more 
longevity:
to publish the code installed in a virtual environment
(including the dependencies used at that time), as described below.
Write code that will stand the test of time perfectly
is deemed 'impossible' \cite{benureau2018re}.
